working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-03-11 Manuscript_draft/Manuscript_draft.Rmd
edits

\newpage
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, out.width='100%', fig.cap="Phylogenetic tree of the ten species used in the experiment from three different families from bottom to top: Solanaceae, Convolvulaceae and Brassicaceae."}
knitr::include_graphics('images/phylo_image.png')
```

@@

Results of hand cross-pollination, hand self-pollination, natural selfing and apomixis are presented in **Figure 1** (see appendix 1 for table with values). Heterospecific pollen reduced seed set significatively with the 50-50% heterospecific pollen treatments for 65% of the pairwise interactions p<0.05. Moreover, average effect sizes differed across species and across families, see **Figure 2**. Despite some variability in the effect of the different pollen donors per species, in general terms the effect of heterospecific pollen from the distinct nine treatments per species was homogeneous (see **Figure 3**), just for four species out of ten, just one donor did not overlap the confidence intervals with other donors. Therefore, none of the donors had a clear stronger or weaker effect across species. When the donors were groped by family not big differences were seen, just for *S. lycopersicum* the confidence intervals of Brassicaceae and Convolvulaceae did not overlap (see **Figure 4**). In addition, for the 100% hetrospecific pollen treatments we did not find almost seed production. However, for just one species (*S. alba*) the control pollination and the heterospecific pollination with pollen from a confamilial had similar seed production. For both Solanaceae species *S. melongena* and *C. annuum* 100% pollen treatments produced few seedles fruits (3% and 9% respectively) and they did not for the apomictic treatments.

Results of Mantel test between heterospecific pollen effect and phylogenetic distance gave a positive statistically clear correlation for both markers (p<0.05). The correlations with ITS and RBCL markers was respectively of 0.29 and 0.25. The phylogenetic signal of traits was significantly correlated with pollen size, stigma measurements and style length (p<0.05). Although with a lack of a significant correlation pagel's lambda values were also relatively high (>0.45) for incompatibility index, ovary length and levels of selfing. Moreover, Mantel test between heterospecific pollen effect and traits gave also a positive significant correlation with a  r value of 0.4. When the effect was look trait by trait with Mantel test, stigma type and stigma measurements (length, width and area) gave a significant positive correlation with heterospecific pollen effect (LOOK AT HOW I EXPRESS THIS IS VERY RELEVANT, HP EFFECT FOR ME IS THE OPPOSITE)

>>>>

\newpage
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, out.width='100%', fig.cap="Phylogenetic tree of the ten species used in the experiment from three different families from top to bottom: Brassicaceae, Convolvulaceae and Solanaceae"}
knitr::include_graphics('images/phylo_image.png')
```

@@

Results of hand cross-pollination, hand self-pollination, natural selfing and apomixis are presented in **Figure 1** (see appendix 1 for table with values). Heterospecific pollen reduced seed set significatively with the 50-50% heterospecific pollen treatments for 65% of the pairwise interactions p<0.05. Moreover, average effect sizes differed across species and across families, see **Figure 2**. Despite some variability in the effect of the different pollen donors per species, in general terms the effect of heterospecific pollen from the distinct nine treatments per species was homogeneous (see **Figure 3**), just for four species out of ten, just one donor did not overlap the confidence intervals with other donors. Therefore, none of the donors had a clear stronger or weaker effect across species. When the donors were groped by family not big differences were seen, just for *S. lycopersicum* the confidence intervals of Brassicaceae and Convolvulaceae did not overlap (see **Figure 4**). In addition, for the 100% hetrospecific pollen treatments we did not find almost seed production. However, for just one species (*S. alba*) the control pollination and the heterospecific pollination with pollen from a confamilial had similar seed production. For two Solanaceae species *S. melongena* and *C. annuum* 100% pollen treatments produced few seedles fruits (3% and 9% respectively) and they did not for the apomictic treatments.

Results of Mantel test between heterospecific pollen effect and phylogenetic distance gave a positive statistically clear correlation for both markers (p<0.05). The correlations with ITS and RBCL markers was respectively of 0.29 and 0.25. We found a significant phylogenetic signal of traits for pollen size, stigma measurements and style length (p<0.05). Although with a lack of a significant correlation pagel's lambda values were also relatively high (>0.45) for incompatibility index, ovary length and levels of selfing. Moreover, Mantel test between heterospecific pollen effect and traits gave also a positive significant correlation with a  r value of 0.4. When the effect was look trait by trait with Mantel test, stigma type and stigma measurements (length, width and area) gave a significant positive correlation with heterospecific pollen effect.
